---
title: "Penguins Size Comparison"
author: "Kat Cotti"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests (compare means between two different groups)

Mass comparision: male vs. female chinstrap penguins (palmer penguins data attached above)
```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap") #Only keep chinstrap pengins

#Histograms - only give one variable bc a histogram creates counts of that variable 
ggplot(data=chinstraps, aes(x=-body_mass_g))  + 
  geom_histogram(bins = 10) +
  facet_wrap(~sex) #splits into two histograms based on sex

#QQ Plot: closer to linear --> closer to normal distribution; give one variable
ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() + 
  facet_wrap(~sex)

#Get Descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
``` 

Get vectors of body mass for female and male chinstrap penguins separately 

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

## T-tests, a couple different ways 

1. Is there a significant difference in body mass between male and female chinstrap penguins? - two-sided question, not asking is one bigger than the other but asking if the mean differnece is large enough that we think they are from population has different means

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest #Calls the stored object so it appears in the markdown document 

cohen.d(chinstraps_f, chinstraps_m) #Gives you the effect size, needs two vectors of values 
```
- Welch's gives --> 
  - Actual Mean values from each group. 
  - Ho = means are equal, Ha = means are NOT equal to 0
  - T-value: how different from 0 the means are on the t-distribution
  -df = measure of sample size
  - p-value: probability that I could have gotten sample means at least as different as ours are if they were drawn from populations with the same mean 
- Cohen.d --> magnitude of effect size -->  metric to decribe the difference between groups that is not so highly impacted by sample size

**SEE ANSWER KEY to coding lesson --> gives info on how to use Broom package to call results of t-tests into nicely organized data frames in the document

#  Are male chinstrap penguins significantly larger in body mass than female chinstrap penguins? --> one sided test, only interested in one side of the test... 

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater") #greater means first value in vector (male) is GREATER than second value in vector (female)
```
- Male chinstrap penguin body mass is significantly different than female chinstrap body mass given the p-value < alpha 

## One-way ANOVA - way to compare means across more than two groups (extension of 2 sample t-tests), have a single variable with discrete groups within it

### 1. Table of counts by species and sex: look at counts per group --> CLT to determine normal distrbution

```{r}
penguins_counts <- penguins %>% 
  count(species, sex) #gives the counts for each variable
penguins_counts
```

Create a subset of only female penguins:
```{r}
penguins_f <- penguins %>% 
  filter(sex=="female") #Create a subset of data for only female penguins

#Make a histogram plot: 
ggplot(data=penguins_f, aes(x = flipper_length_mm)) + #give one variable --> counts of the var
  geom_histogram(aes(fill = species)) + #changes color of the different species
  facet_wrap(~species, scales = "free") #splits the hisogram into 3 different based on species, facet wrap default is that axes labels have to match --> if you want to look at distribution within groups using scales = "free" --> see distribution with axes based on each group 

#Make a qq Plot: closer to linear --> closer to normally distributed population 
ggplot(data=penguins_f, aes(sample = flipper_length_mm)) + #give one sample variable
  geom_qq(aes(color = species)) +
  facet_wrap(~species, scales = "free")
```

### Find summary statistics:

starting from penguins_f. create a summary table that groups by species and returnd the following
- mean of flipper length
- sd of flipper length
- sample size of flipper length
- variance 'var'
- standard error (sd of the sampling distribution) = sd/sqrt(n)

```{r}
penguins_f_summary <- penguins_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    var_flipper_length = var(flipper_length_mm)
    )
penguins_f_summary

#As sample size increases you WILL detect differences in variance - same flaw as hypothesis testing - hypothesis test with null that groups have = variance
leveneTest(flipper_length_mm ~ species, data = penguins_f)
```
p-value above alpha --> retain null hpothesis that the variances are equal 

### Make a Beeswarm plot with the mean +/- one standard deviation - need beeswarm package attached - gives actual values of observations and arranges horizontally to give clear picture of where greater densitites of observations are

```{r}
ggplot() +
  geom_beeswarm(data = penguins_f,#plots actual data
                aes(x = species, y = flipper_length_mm), #in aes refers to a variable 
                color = "gray70", #changes color of the points 
                size = 3, #changes size of the points
                alpha = 0.5) +   #changes transparency
  geom_point(data = penguins_f_summary, aes(x = species,#ggplot works progressively so this will be ON top on the beeswarm
                                            y = mean_flipper_length), 
                                             size = 4) +
  geom_errorbar(data = penguins_f_summary, #adds whiskers - shows +/- one standard deviation 
                aes(x = species,
                    ymax = mean_flipper_length +sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) +
  theme_light()


```
### One-Way ANOVA test - aov function runs a one way ombibus ANOVA, usually we will want to run a type 3 anova test;
- omnibus ANOVA allows us to reject that all means are equal -> if reject run a post hoc test

```{r}
penguin_aov <- aov(flipper_length_mm ~species, data = penguins_f)
summary(penguin_aov) #look at the outcome

#Post-hoc test to run pairwise comparisons: gives p values for pairwise adjusted values --> all are below 0.05 means there is a significant difference between ALL means 
TukeyHSD(penguin_aov)
```
-F stat = ratio of the Between group variance / within group sum of squares -> large value means Between groups variance >> within group variance 
p-value = probability that we could have gotten sample means at least this different if the samples were drawn from populations with the same mean --> very very small P value = reject the null hypothesis that the groups are from populations with the same mean --> all means are not the same!


### Type III ANOVA with the car package:
IF you have very unbalanced design (large diff in sample size, distributions, ect.) --> Use Type 3 or Orthoganal ANOVA test:
```{r}
#Create a linear model:
penguin_lm <- lm(flipper_length_mm ~species, data = penguins_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```
F-Value  - most of the time it will be very close to the F value from type 1 unless data is super super unbalanced 


